---
title: "Community-Structured Brain Network Populations"
author: "Maximiliano Martino, Daniel Fraiman"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Community-Structured Brain Network Populations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Purpose

This vignette demonstrates a complete three-group workflow on synthetic
stochastic block model networks. The simulation is an API illustration, not
evidence about a clinical population.

```{r}
library("BrainNetTest")
```

## Generating one graph

`generate_community_graph()` supports one within-community probability per
community and a common between-community probability.

```{r}
community_sizes <- c(5, 5)
graph <- generate_community_graph(
  n_nodes = 10,
  n_communities = 2,
  community_sizes = community_sizes,
  intra_prob = c(0.85, 0.65),
  inter_prob = 0.1,
  seed = 1
)

isSymmetric(graph)
all(diag(graph) == 0)
```

## Simulating three independent groups

```{r}
group_a <- generate_category_graphs(
  n_graphs = 10,
  n_nodes = 10,
  n_communities = 2,
  community_sizes = community_sizes,
  base_intra_prob = c(0.85, 0.65),
  base_inter_prob = 0.10,
  seed = 11
)
group_b <- generate_category_graphs(
  n_graphs = 10,
  n_nodes = 10,
  n_communities = 2,
  community_sizes = community_sizes,
  base_intra_prob = c(0.55, 0.65),
  base_inter_prob = 0.25,
  seed = 12
)
group_c <- generate_category_graphs(
  n_graphs = 10,
  n_nodes = 10,
  n_communities = 2,
  community_sizes = community_sizes,
  base_intra_prob = c(0.30, 0.65),
  base_inter_prob = 0.45,
  seed = 13
)

networks <- brainnet_data(
  list(GroupA = group_a, GroupB = group_b, GroupC = group_c),
  node_labels = paste0("Region", seq_len(10)),
  communities = rep(c("Module 1", "Module 2"), community_sizes)
)
summary(networks)
```

## Omnibus analysis

For more than two groups, edge-wise Fisher tests are omnibus tests of equal
edge frequency. `effect_range` reports the largest observed difference among
group proportions, but does not make a pairwise inferential claim.

```{r}
result <- brainnet_test(
  networks,
  n_permutations = 199,
  adjust = "holm",
  seed = 21
)

result
head(as.data.frame(result)[order(as.data.frame(result)$rank), ])
selected_edges(result)
```

The global result and edge results answer related but distinct questions.
Selected edges are determined only by their multiplicity-adjusted edge tests;
they are not labeled as causes of a global rejection.

## Plotting the classed result

The `plot()` method uses central graphs and metadata stored in the result, so
the original populations do not need to be passed again.

```{r plot, eval = requireNamespace("igraph", quietly = TRUE), fig.width=8, fig.height=8}
plot(result, reference = "GroupA")
```

`selected_nodes()` counts incidences of selected edges. This is a descriptive
summary and does not provide node-level error control.

```{r}
selected_nodes(result)
```

## Interpretation limits

The simulation changes marginal edge probabilities directly. Performance in
this example should not be extrapolated to weighted networks, directed
networks, estimated connectomes with preprocessing uncertainty, paired data,
or alternatives that change only dependence between edges.
