---
title: "Introduction to BrainNetTest"
author: "Maximiliano Martino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BrainNetTest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Introduction
BrainNetTest is an R package designed to perform hypothesis testing on brain networks represented as graphs. It provides functionalities to:

Generate random brain networks.
Compute central graphs for populations.
Calculate distances between graphs.
Compute test statistics to assess differences between populations.
This vignette introduces the basic usage of BrainNetTest.

Installation
You can install the development version of BrainNetTest from GitHub:

```{r setup, include = FALSE}
# Install devtools if not already installed
install.packages("devtools")

# Install BrainNetTest from GitHub
devtools::install_github("mmaximiliano/BrainNetTest")
```

Getting Started
Load the BrainNetTest package:

```{r setup, include = FALSE}
library(BrainNetTest)
```

Generating Random Brain Networks
You can generate random symmetric adjacency matrices representing brain networks using the generate_random_graph function.

```{r setup, include = FALSE}
# Generate a random brain network with 10 nodes and edge probability of 0.1
G <- generate_random_graph(n_nodes = 10, edge_prob = 0.1)
print(G)
```

Computing the Central Graph
The compute_central_graph function computes the central (representative) graph for a population.

```{r setup, include = FALSE}
# Generate synthetic data for a population
population <- list(
  generate_random_graph(n_nodes = 10, edge_prob = 0.1),
  generate_random_graph(n_nodes = 10, edge_prob = 0.1)
)

# Compute the central graph
central_graph <- compute_central_graph(population)
print(central_graph)
```

Computing Distances Between Graphs
Use compute_distance to calculate the Manhattan norm distance between two graphs.

```{r setup, include = FALSE}
# Compute distance between a graph and the central graph
distance <- compute_distance(population[[1]], central_graph)
print(distance)
```

Computing the Test Statistic
The compute_test_statistic function computes the test statistic T to assess whether different populations of brain networks originate from the same distribution.

```{r setup, include = FALSE}
# Generate synthetic data for three populations
Control <- replicate(30, generate_random_graph(n_nodes = 10, edge_prob = 0.1), simplify = FALSE)
Schizophrenia <- replicate(30, generate_random_graph(n_nodes = 10, edge_prob = 0.15), simplify = FALSE)
Alzheimer <- replicate(30, generate_random_graph(n_nodes = 10, edge_prob = 0.2), simplify = FALSE)

# Combine populations into a list
populations <- list(Control = Control, Schizophrenia = Schizophrenia, Alzheimer = Alzheimer)

# Compute the test statistic T
T_value <- compute_test_statistic(populations, a = 1)
print(T_value)
```

Conclusion
This vignette provided an introduction to the basic functionalities of BrainNetTest. In the next vignette, we'll explore generating brain networks with community structures and analyzing them.

References
