---
title: "Generating and Analyzing Brain Networks with Community Structures"
author: "Maximiliano Martino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Brain Networks with Community Structures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Introduction
In this vignette, we'll explore how to generate brain networks with community structures using BrainNetTest and analyze them to perform hypothesis testing.

Generating Brain Networks with Community Structures
Using generate_community_graph
The generate_community_graph function generates a single brain network with specified community structures.

```{r setup, include = FALSE}
library(BrainNetTest)

# Generate a brain network with community structure
G <- generate_community_graph(
  n_nodes = 100,
  n_communities = 4,
  intra_prob = 0.8,
  inter_prob = 0.2,
  seed = 42
)

# Display a portion of the adjacency matrix
print(G[1:10, 1:10])
```

Visualizing the Network
We can visualize the network to observe the community structure.

```{r setup, include = FALSE}
# Install and load igraph for visualization
if (!requireNamespace("igraph", quietly = TRUE)) {
  install.packages("igraph")
}
library(igraph)

# Create an igraph object
graph <- graph_from_adjacency_matrix(G, mode = "undirected")

# Assign community membership as vertex attributes
community_sizes <- NULL
n_communities <- 4
n_nodes <- 100
if (is.null(community_sizes)) {
  base_size <- n_nodes %/% n_communities
  remainder <- n_nodes %% n_communities
  community_sizes <- rep(base_size, n_communities)
  if (remainder > 0) {
    community_sizes[1:remainder] <- community_sizes[1:remainder] + 1
  }
}
community_assignments <- rep(1:n_communities, times = community_sizes)
V(graph)$community <- community_assignments

# Plot the network
set.seed(123)
plot(graph, vertex.color = V(graph)$community, vertex.size = 5, vertex.label = NA)
```

Generating Graphs for a Category with Similar Community Structures
Using generate_category_graphs
The generate_category_graphs function generates a set of graphs belonging to the same category with similar community structures.

```{r setup, include = FALSE}
# Generate 5 graphs for the Control category
control_graphs <- generate_category_graphs(
  n_graphs = 5,
  n_nodes = 100,
  n_communities = 4,
  base_intra_prob = 0.8,
  base_inter_prob = 0.2,
  intra_prob_variation = 0.05,
  inter_prob_variation = 0.05,
  seed = 1
)

# Check the number of graphs generated
length(control_graphs)
```

Performing Hypothesis Testing Between Categories
Generating Graphs for Multiple Categories

```{r setup, include = FALSE}
# Generate graphs for the Control category
control_graphs <- generate_category_graphs(
  n_graphs = 30,
  n_nodes = 100,
  n_communities = 4,
  base_intra_prob = 0.8,
  base_inter_prob = 0.2,
  intra_prob_variation = 0.05,
  inter_prob_variation = 0.05,
  seed = 1
)

# Generate graphs for the Disease category
disease_graphs <- generate_category_graphs(
  n_graphs = 30,
  n_nodes = 100,
  n_communities = 4,
  base_intra_prob = 0.6,
  base_inter_prob = 0.4,
  intra_prob_variation = 0.05,
  inter_prob_variation = 0.05,
  seed = 2
)

# Combine populations into a list
populations <- list(Control = control_graphs, Disease = disease_graphs)
```

Computing the Test Statistic

```{r setup, include = FALSE}
# Compute the test statistic T
T_value <- compute_test_statistic(populations, a = 1)
print(T_value)
```

Interpreting the Results
A higher value of the test statistic T indicates greater differences between the populations' brain networks.

Conclusion
This vignette demonstrated how to generate brain networks with community structures and perform hypothesis testing using BrainNetTest. By simulating realistic brain network data, researchers can test hypotheses and analyze differences between various categories, such as healthy controls and patients with specific conditions.

References