---
title: "Introduction to BrainNetTest"
author: "Maximiliano Martino, Daniel Fraiman"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BrainNetTest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup_intro, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Scope

`BrainNetTest` compares independent groups of subject-level binary,
undirected brain networks observed on the same labeled atlas. Typical inputs
are thresholded functional or structural connectomes, one per participant.
The package deliberately separates:

* a whole-graph label-randomization test, whose score is targeted to marginal
  edge-frequency shifts; and
* edge-wise tests with multiplicity adjustment.

The global test is not omnibus for alternatives that preserve every edge's
marginal probability while changing only dependence, motifs, or other
topology. Weighted, directed, paired, clustered, and covariate-adjusted designs
are outside the current contract.

Atlas selection, image registration, connectivity estimation, nuisance
correction, thresholding, and quality control are upstream analysis choices.
They are not performed by `BrainNetTest`, and their uncertainty must be
considered when results are interpreted. Selected edges describe statistical
group associations; they are not automatically causal links or clinical
biomarkers.

## Constructing validated data

```{r}
library("BrainNetTest")

group_a <- generate_category_graphs(
  n_graphs = 12,
  n_nodes = 8,
  n_communities = 2,
  base_intra_prob = 0.8,
  base_inter_prob = 0.1,
  seed = 1
)
group_b <- generate_category_graphs(
  n_graphs = 12,
  n_nodes = 8,
  n_communities = 2,
  base_intra_prob = 0.4,
  base_inter_prob = 0.5,
  seed = 2
)

networks <- brainnet_data(
  list(GroupA = group_a, GroupB = group_b),
  node_labels = paste0("Node", seq_len(8)),
  communities = rep(c("Community 1", "Community 2"), each = 4)
)
networks
summary(networks)
```

The constructor rejects missing values, weighted or asymmetric matrices,
self-loops, inconsistent dimensions, and inconsistent node ordering before
any inferential computation begins.

## Graph summaries

`central_graph()` is the element-wise average of a population. For undirected
graphs, `graph_distance()` counts every upper-triangle edge exactly once.

```{r}
center_a <- central_graph(group_a)
graph_distance(group_a[[1]], center_a)
```

## Global and edge-wise inference

```{r}
result <- brainnet_test(
  networks,
  n_permutations = 199,
  edge_method = "fisher",
  adjust = "holm",
  seed = 42
)

result
summary(result)
```

The global p-value uses inclusive ties and the Monte Carlo plus-one correction.
When the number of fixed-size label assignments is at most 10,000,
`exact = "auto"` exhaustively enumerates them.

The edge table is a separate family of hypotheses. It is not a decomposition
of the global result and is not gated on global rejection.

```{r}
head(as.data.frame(result))
selected_edges(result)
selected_nodes(result)
```

Holm adjustment is the default because it controls family-wise error under
arbitrary dependence among edge tests. BY is available for false-discovery
rate control under arbitrary dependence. BH is opt-in because it requires
additional positive-dependence assumptions.

## Exploratory ablation

```{r}
result_with_ablation <- brainnet_test(
  networks,
  n_permutations = 99,
  ablation = TRUE,
  seed = 42
)
head(ablation_path(result_with_ablation))
```

The ablation ordering is selected from the same data. Consequently,
`descriptive_tail_fraction` is not a post-selection p-value, and the path does
not establish equivalence, minimality, necessity, or causality.

## Reference

Fraiman D and Fraiman R (2018). "An ANOVA approach for statistical comparisons
of brain networks." *Scientific Reports*, **8**, 4746.
<https://doi.org/10.1038/s41598-018-23152-5>
