---
title: "Introduction to BrainNetTest"
author: "Maximiliano Martino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BrainNetTest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup_intro, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

## Introduction

**BrainNetTest** is a comprehensive R package designed for statistical analysis and hypothesis testing of brain network populations. In computational neuroscience, brain networks are commonly represented as undirected graphs where nodes correspond to brain regions (e.g., parcellated areas from neuroimaging) and edges represent functional or structural connections between these regions.

This package addresses the critical need for robust statistical methods to:

1. **Detect population-level differences** between groups (e.g., healthy controls vs. patients)
2. **Identify specific connections** that drive observed differences
3. **Perform rigorous statistical inference** at both global and edge levels
4. **Generate realistic synthetic networks** for method validation

## Key Concepts

### Brain Networks as Graphs

In BrainNetTest, each brain network is represented as:
- **Adjacency matrix**: Symmetric binary matrix where 1 indicates connection, 0 indicates no connection
- **No self-loops**: Diagonal elements are always 0
- **Undirected**: Matrix is symmetric (connections are bidirectional)

### Population Analysis

The package works with **populations** of brain networks:
- Each population represents a group (e.g., "Control", "Patient") 
- Each group contains multiple individual brain networks
- Statistical tests compare distributions across populations

## Installation

Install the development version from GitHub:

```{r installation, eval = FALSE}
# Install remotes if needed
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}

# Install BrainNetTest
remotes::install_github("mmaximiliano/BrainNetTest", build_vignettes = TRUE)
```

## Getting Started

Load the package:

```{r load_package}
library(BrainNetTest)
```

## Core Functionalities

### 1. Generating Synthetic Brain Networks

BrainNetTest provides tools to generate realistic synthetic brain networks for testing and validation:

```{r generate_networks}
# Set seed for reproducibility
set.seed(42)

# Generate a simple random network
simple_network <- generate_random_graph(n_nodes = 20, edge_prob = 0.15)

# Display network properties
cat("Network size:", nrow(simple_network), "x", ncol(simple_network), "\n")
cat("Number of edges:", sum(simple_network) / 2, "\n")
cat("Density:", round(sum(simple_network) / (nrow(simple_network) * (nrow(simple_network) - 1)), 3), "\n")

# Show a portion of the adjacency matrix
print(simple_network[1:8, 1:8])
```

### 2. Generating Networks with Community Structure

Real brain networks exhibit community structure (modules). Generate more realistic networks:

```{r community_networks}
# Generate networks with community structure
community_networks <- generate_category_graphs(
  n_graphs = 5,
  n_nodes = 50,
  n_communities = 4,
  base_intra_prob = 0.6,    # High connectivity within communities
  base_inter_prob = 0.1,    # Low connectivity between communities
  seed = 123
)

cat("Generated", length(community_networks), "networks\n")
cat("Each network has", nrow(community_networks[[1]]), "nodes\n")

# Compare densities
densities <- sapply(community_networks, function(g) sum(g) / (nrow(g) * (nrow(g) - 1)))
cat("Network densities:", round(densities, 3), "\n")
```

### 3. Computing Central (Representative) Graphs

For each population, compute a central graph that represents the typical connectivity pattern:

```{r central_graphs}
# Generate two populations with different properties
control_population <- generate_category_graphs(
  n_graphs = 20,
  n_nodes = 30,
  base_intra_prob = 0.7,
  base_inter_prob = 0.1,
  seed = 456
)

patient_population <- generate_category_graphs(
  n_graphs = 20,
  n_nodes = 30,
  base_intra_prob = 0.5,    # Reduced intra-community connectivity
  base_inter_prob = 0.2,    # Increased inter-community connectivity
  seed = 789
)

# Compute central graphs
control_central <- compute_central_graph(control_population)
patient_central <- compute_central_graph(patient_population)

cat("Central graphs computed\n")
cat("Control central graph density:", 
    round(sum(control_central) / (nrow(control_central) * (nrow(control_central) - 1)), 3), "\n")
cat("Patient central graph density:", 
    round(sum(patient_central) / (nrow(patient_central) * (nrow(patient_central) - 1)), 3), "\n")
```

### 4. Computing Distances Between Networks

Measure how similar/different networks are using Manhattan distance:

```{r distances}
# Compute distances from individuals to their population centrals
control_distances <- sapply(control_population, function(g) compute_distance(g, control_central))
patient_distances <- sapply(patient_population, function(g) compute_distance(g, patient_central))

cat("Control population distances (first 5):", round(control_distances[1:5], 2), "\n")
cat("Patient population distances (first 5):", round(patient_distances[1:5], 2), "\n")

# Summary statistics
cat("\nDistance summary:\n")
cat("Control - Mean:", round(mean(control_distances), 2), 
    ", SD:", round(sd(control_distances), 2), "\n")
cat("Patient - Mean:", round(mean(patient_distances), 2), 
    ", SD:", round(sd(patient_distances), 2), "\n")
```

### 5. Global Population Testing

Test whether populations have significantly different network properties:

```{r global_test}
# Combine populations for testing
populations <- list(Control = control_population, Patient = patient_population)

# Compute test statistic
T_statistic <- compute_test_statistic(populations, a = 1)

cat("T-statistic:", round(T_statistic, 4), "\n")
cat("Interpretation: Larger |T| values indicate greater population differences\n")
```

### 6. Edge-Level Analysis

Identify which specific connections differ between populations:

```{r edge_analysis}
# Compute edge frequencies across populations
edge_frequencies <- compute_edge_frequencies(populations)

cat("Edge frequency analysis:\n")
cat("Dimensions of edge counts:", dim(edge_frequencies$edge_counts), "\n")

# Compute p-values for each edge
population_sizes <- sapply(populations, length)
edge_pvalues <- compute_edge_pvalues(
  edge_frequencies$edge_counts, 
  population_sizes,
  method = "fisher",
  adjust_method = "fdr"
)

# Rank edges by significance
ranked_edges <- rank_edges(edge_pvalues)

cat("Number of edges tested:", nrow(ranked_edges), "\n")
cat("Top 5 most significant edges:\n")
print(head(ranked_edges, 5))
```

### 7. Identifying Critical Connections

Find the specific edges that explain population differences:

```{r critical_edges}
# Identify critical links
critical_results <- identify_critical_links(
  populations = populations,
  alpha = 0.05,
  method = "fisher",
  n_bootstrap = 500,  # Reduced for vignette speed
  seed = 999
)

if (!is.null(critical_results$critical_edges)) {
  cat("Number of critical edges identified:", nrow(critical_results$critical_edges), "\n")
  cat("Critical edges:\n")
  print(head(critical_results$critical_edges))
} else {
  cat("No critical edges identified at α = 0.05\n")
}

cat("Number of edges physically removed:", length(critical_results$edges_removed), "\n")
```

## Practical Example: Simulated Clinical Study

Let's simulate a realistic clinical study comparing brain networks:

```{r clinical_example}
# Simulate a clinical study with larger populations
set.seed(2024)

# Healthy controls: Strong community structure
healthy_nets <- generate_category_graphs(
  n_graphs = 50,
  n_nodes = 80,
  n_communities = 5,
  base_intra_prob = 0.75,
  base_inter_prob = 0.05,
  intra_prob_variation = 0.1,
  inter_prob_variation = 0.02
)

# Patients: Disrupted community structure
patient_nets <- generate_category_graphs(
  n_graphs = 45,
  n_nodes = 80,
  n_communities = 5,
  base_intra_prob = 0.45,    # Much weaker communities
  base_inter_prob = 0.15,    # More random connections
  intra_prob_variation = 0.15,
  inter_prob_variation = 0.05
)

study_populations <- list(Healthy = healthy_nets, Patient = patient_nets)

# Global test
study_T <- compute_test_statistic(study_populations, a = 1)
cat("Study T-statistic:", round(study_T, 3), "\n")

# Edge-level analysis with multiple testing correction
study_frequencies <- compute_edge_frequencies(study_populations)
study_sizes <- sapply(study_populations, length)

study_pvalues <- compute_edge_pvalues(
  study_frequencies$edge_counts,
  study_sizes,
  method = "fisher",
  adjust_method = "bonferroni"
)

study_edges <- rank_edges(study_pvalues)
significant_edges <- study_edges[study_edges$p_adjusted < 0.05, ]

cat("Significant edges (Bonferroni corrected):", nrow(significant_edges), "\n")
if (nrow(significant_edges) > 0) {
  cat("Most significant edge: nodes", significant_edges$node1[1], "-", 
      significant_edges$node2[1], ", p =", format(significant_edges$p_adjusted[1], scientific = TRUE), "\n")
}
```

## Best Practices

### Sample Size Considerations
- **Minimum**: 20-30 networks per population for reliable statistics
- **Recommended**: 50+ networks for robust edge-level inference
- **Power**: Larger samples detect smaller effect sizes

### Statistical Considerations
- Use **multiple testing correction** for edge-level analysis
- Consider **effect sizes** not just p-values
- **Bootstrap iterations**: 1000+ for critical link identification

### Computational Efficiency
- Start with smaller networks for exploratory analysis
- Use `batch_size > 1` in `identify_critical_links()` for faster computation
- Set `seed` parameters for reproducible results

## Next Steps

This vignette introduced the core functionality of BrainNetTest. For more advanced topics:

- **Community Structure Analysis**: See the "Community Structures" vignette
- **Function Documentation**: Use `?function_name` for detailed parameter descriptions
- **GitHub Repository**: Visit for latest updates and examples

## References

For methodological details and applications, refer to the package documentation and associated publications.
